### 2018-06-18 mmm-minorcycle.tex
Rev 1.0

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Distribution of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-CASS-001\xspace}
% Context
\newcommand{\context}{(ASKAPsoft)}
% Revision
\newcommand{\revision}{0.9\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{Draft for review\xspace}

@@

\newcommand{\secondsigname}{T.J. Cornwell}
\newcommand{\secondsigdesignation}{Lead Consultant}
\newcommand{\secondsigaffiliation}{TCC}

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
    0.9 & 30/04/2018 & T.J. Cornwell & Initial version \\
    \hline
  \end{tabularx}
}

@@

In the current single threaded \MAM\ the minor cycles, the processing consists of a search step, scaling as the number of pixels, and a subtraction step,  scaling as the product of the number of pixels in the image and the size of the PSF support. The value of the PSF support used thus directly affects the processing time. Furthermore, the processing via sub-image inevitably means that a limit PSF support will be used.

@@

Clearly, in the context of a major/minor cycle algorithm, small PSF support can be used: the speedup is significant and there is no significant increase in error compared to the cutoff for the transition from minor to major cycle. This is confirmed by evaluating the difference between images processed using \MAM\ and Sub-image \MAM\ after ten major cycles (see figure \ref{fig:error_vs_support_ci}.

We should emphasize that this result also holds for the current use of the single threaded \MAM\: the psf support can be set to a value much smaller than 512. One possible limit would be where the subtraction step becomes comparable to the search cost. This is obviously context dependent but we would expect values in the range 32 - 128. However, figure \ref{fig:time_vs_support} shows that the gains are relatively small.

@@

The next plot (figure \ref{fig:fractions_with_pb}) shows the effect of the ASKAP (single) primary beam. Most sub-images do not require cleaning and all the processing goes to the brightest fields. As processing proceeds thorough a number of major cycles, the brightest sources are been removed and eventually there remains a sea of sources spread other the primary and into the primary beam side-lobes. For the final major cycles, we can model the flux as simply by clipping the image with the primary beam at some maximum brightness such as 1mJy. The resulting scaling shown in figure \ref{fig:fractions_clipped_image} is then better balanced.

So, to summarise, we expect the scaling to be poor for the initial major cycles (typically $\sim 5$ for 16 facets per axis) where the brightest sources in the main lobes are removed, but it will improve (typically to $\sim 50$) as the cleaning goes deeper. Since the number of weak sources is much greater we can expect the scaling to be close to that on figure \ref{fig:fractions_clipped_image}

@@

Since the scaling curves are relatively flat (see figures \ref{fig:aggregate_scaling} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing. Load balancing would then only be needed for the high rank tail and should be performed automatically by the master/worker pattern.

@@

To facilitate development, the sub-image \MAM\ should first be implemented in cdeconvolver to resolve the issues around use of the master-worker scheme and then ported to cimager.

>>>>

% Enter here some information needed to fill in the template Title to appear
% on the front pages (will be filled in via the \sdpfrontpage command)
\newcommand{\bigdoctitle}{Distribution of the Rau-Cornwell MSMFS algorithm \xspace}
% Title to go in the "Document Status Sheet" Document number
\newcommand{\docnr}{TCC-CASS-001\xspace}
% Context
\newcommand{\context}{(ASKAPsoft)}
% Revision
\newcommand{\revision}{0.9\xspace}
% Author(s)
\newcommand{\docauthor}{T.J.\ Cornwell\xspace}
% Lead author (goes in the footer)
\newcommand{\leadauthor}{T.J.\ Cornwell\xspace}
% Release date (YYYY-MM-DD)
\newcommand{\docudate}{\today\ \currenttime}
% Document classification
\newcommand{\classification}{Unrestricted}
% Status of the document (draft/final/etc.)
\newcommand{\docstatus}{For release\xspace}

@@

% Table with version numbers
\newcommand{\versiontable}{
  \begin{tabularx}{\textwidth}{|X|X|X|X|}
    \hline
    \bf{Version} & {\bf Date of issue} & {\bf Prepared by} & {\bf Comments}\\
    \hline
    0.9 & 30/04/2018 & T.J. Cornwell & Initial version \\
    \hline
    1.0 & 18/06/2018 & T.J. Cornwell & Corrections from Mitch \\
    \hline
  \end{tabularx}
}

@@

In the current single threaded \MAM\ the minor cycles, the processing consists of a search step, scaling as the number of pixels, and a subtraction step,  scaling as the square of the size of the PSF support. The value of the PSF support used thus directly affects the processing time. Furthermore, the processing via sub-image inevitably means that a limit PSF support will be used.

@@

Clearly, in the context of a major/minor cycle algorithm, small PSF support can be used: the speedup is significant and there is no significant increase in error compared to the cutoff for the transition from minor to major cycle. This is confirmed by evaluating the difference between images processed using \MAM\ and Sub-image \MAM\ after ten major cycles (see figure \ref{fig:error_vs_support_ci}).

We should emphasize that this result also holds for the current use of the single threaded \MAM\ : the psf support can be set to a value much smaller than 512. One possible limit would be where the subtraction step becomes comparable to the search cost. This is obviously context dependent but we would expect values in the range 32 - 128. However, figure \ref{fig:time_vs_support} shows that the gains are relatively small.

@@

The next plot (figure \ref{fig:fluxes_with_pb}) shows the effect of the ASKAP (single) primary beam. Most sub-images do not require cleaning and all the processing goes to the brightest fields. As processing proceeds thorough a number of major cycles, the brightest sources are been removed and eventually there remains a sea of sources spread other the primary and into the primary beam side-lobes. For the final major cycles, we can model the flux as simply by clipping the image with the primary beam at some maximum brightness such as 1mJy. The resulting scaling shown in figure \ref{fig:fractions_clipped_image} is then better balanced.

So, to summarise, we expect the scaling to be poor for the initial major cycles (typically $\sim 5$ for 16 facets per axis) where the brightest sources in the main lobes are removed, but it will improve (typically to $\sim 50$) as the cleaning goes deeper. Since the number of weak sources is much greater we can expect the scaling to be close to that on figure \ref{fig:fractions_clipped_image}.

@@

Since the scaling curves are relatively flat (see figures \ref{fig:aggregate_scaling_2km} and \ref{fig:aggregate_scaling_6km}), most processors will be kept close to fully busy while the most demanding (rank 0) case will be processing. Load balancing would then only be needed for the high rank tail and should be performed automatically by the master/worker pattern.

@@

The gather/scatter step could be implemented either centrally or via neighbourhoods of workers. The latter avoids the use of a large amount of memory on the master to store the entire image.

To facilitate development, the sub-image \MAM\ should first be implemented in cdeconvolver to resolve the issues around use of the master-worker scheme and then ported to cimager. The implementation should accommodate both central and localised gather/scatter as described in the previous paragraph.
